---
title:      "Practical Machine Learning: Peer Assessment"
author:     "Haidong Gu"
date:       "22, March, 2014"
output: 
  html_document:
    keep_md: true
---

## Requirement
please see [Coursera Website](https://class.coursera.org/predmachlearn-012/human_grading/view/courses/973547/assessments/4/submissions) for details.

## Loading and Processing

load librarires
```{r}
library(caret)
library(randomForest)

set.seed(8880)
```

set working directory, it will be different for each person
```{r}
setwd("/Users/hgu/Downloads/")
paste("working dir is", getwd())
```

 1. Load the data

```{r}
if (!file.exists("./pml-training.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
        destfile = "./pml-training.csv")
}
if (!file.exists("./pml-testing.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
        destfile = "./pml-testing.csv")
}

data_training <- read.csv("./pml-training.csv")
data_testing <- read.csv("./pml-testing.csv")
```

 2. Process/transform the data

remove columns with all missing values or without keywords from predication.
```{r}
missing_value_col <- sapply(data_testing, function (x) any(is.na(x) | x == ""))
predictors <- names(missing_value_col)[!missing_value_col]
predictors <- predictors[grepl("belt|forearm|arm|dumbbell", predictors)]

data_training <- data_training[c("classe", predictors)]

dim(data_training)
names(data_training)
```

 3. partition the training data with 60% training, 40% probing.

```{r}
partition <- createDataPartition(data_training$classe, p=0.6, list=FALSE)
data_probing <- data_training[-partition, ]
data_training <- data_training[partition, ]

```

## Modeling

building model with Random Forest ("rf") machine leaning techique. __It will be a good model  if the error is less than 2.5%.__

```{r cache=TRUE}
model <- train(classe ~ ., data = data_training, method = "rf", prox = TRUE, 
               trControl = trainControl(method = "cv", number = 4, allowParallel = TRUE))
```

### Evaluate the model on the training data

```{r}
training_result <- predict(model, data_training)
confusionMatrix(training_result, data_training$classe)
```

### Evaluate the model on the probing data

```{r}
probing_result <- predict(model, data_probing)
confusionMatrix(probing_result, data_probing$classe)
```

### Final Mode
```{r}
varImp(model)
model$finalModel
```

__OOB estimate of error rate is 0.88%.__

## Apply model to testing data and Submit
```{r}
testing_result <- predict(model, data_testing)
```

prepare for submission
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=file.path("/Users/hgu/Downloads/answers2", filename),quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(testing_result)
```

## Summary
* I use _Random Forest_ to to build our model.
* I was expecting error rate to below 2.5%, and my model achieves 0.88%.